### 2017-03-17 paper/main.tex
Initialized project structure

\documentclass[12pt]{article}
\usepackage{amsmath}

\begin{document}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\end{document}

### 2017-03-24 project_plan/apt-project.tex
Project plan added

\documentclass[times, utf8]{article}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{listings}
\usepackage{xcolor}
\newcommand\myworries[1]{\textcolor{red}{#1}}
\usepackage{hyperref} 
\usepackage{natbib}
\usepackage{subfig}
\usepackage{color}
\usepackage{placeins}
\usepackage[T1]{fontenc}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color
  basicstyle=\footnotesize,        % size of fonts used for the code
  breaklines=true,                 % automatic line breaking only at whitespace
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  keywordstyle=\color{blue},       % keyword style
  stringstyle=\color{mymauve},     % string literal style
}

\begin{document}

\title{\textbf{Semantic textual similarity}}
\date{March, 2017}
\author{Bruno Gavranovi\'c, Neven Miculini\'c and Stipan Mikuli\'c}
\maketitle

\section*{Topic description}

Semantic textual similarity (STS) measures how similar two text snippets are. The task is, given two sentences, to determine their similarity score on a continuous scale from 0 to 5. The score of 5 denotes a perfect semantic equivalence. This task was offered in multiple tracks, covering various cross-lingual and monolingual pairs. Your task, however, is the track concerned with English monolingual pairs. Of course, you are encouraged to give a shot on other tracks as well.
\section*{Project plan}
First and foremost we want to merge together the data from different sources and build vocabulary. We plan to develop simple baseline first and then build upon that.
After reading referenced papers we decided that our pipeline will look something like the following:

\begin{enumerate}
\item Data preprocessing

\begin{itemize}
\item Tokenization
\item POS tagging
\item Parsing (syntax)
\item ...
\end{itemize}

\item Transforming data to corresponding input for models
\begin{itemize}
\item N-gram overlap
\item TF-IDF vector
\item Paragraph2Vec
\end{itemize}

\item Building different models
\begin{itemize}
\item LSTM neural network
\item SVM
\item Linear Regression
\end{itemize}

\item Evaluation
\end{enumerate}

We will make baseline model as centroid vector of word embeddings from sentence.
\end{document}

### 2017-05-29 src/baseline.py
Added stopword removal to cosine baseline

@@ -14,9 +14,9 @@ class SimpleBaseline(Model):
         b = {SimpleBaseline.porter.stem(x) for x in b}
         return len(a & b) / len(a | b)
 
-    def preprocessLine(self, sentence):
+    def preprocess_line(self, sentence):
         return nltk.word_tokenize(sentence)
 
 
 model = SimpleBaseline("log", eval_ds, {})
-model.evalScore()
+model.eval_score()

### 2017-05-29 src/cosine_baseline.py

@@ -10,15 +10,20 @@ from model_utils import Model
 class SimpleBaseline(Model):
     porter = nltk.PorterStemmer()
 
+    def __init__(self, *args, **kwargs):
+        super().__init__(*args, **kwargs)
+        self.stopwords = set(nltk.corpus.stopwords.words("english"))
+
     def _predict(self, a, b):
         fn = np.sum
         sem1 = fn(a, axis=0).reshape(1, -1)
         sem2 = fn(b, axis=0).reshape(1, -1)
         return cosine_similarity(sem1, sem2)[0, 0]
 
-    def preprocessLine(self, sentence):
-        return sentence_to_vector(nltk.word_tokenize(sentence))
+    def preprocess_line(self, sentence):
+        tokens = [token for token in nltk.word_tokenize(sentence) if token not in self.stopwords]
+        return sentence_to_vector(tokens)
 
 
 model = SimpleBaseline("log_cosine", eval_ds, {'log_overwrite': True})
-model.evalScore()
+model.eval_score()

### 2017-05-29 src/data.py

@@ -42,6 +42,7 @@ eval_ds = DataSet(eval_data_in_path, eval_data_out_path, "Eval dataset")
 try:
     word2vec = gensim.models.KeyedVectors.load_word2vec_format(w2v_path, binary=False, limit=10000)
 
+
     def sentence_to_vector(sentence):
         """
 

### 2017-05-29 src/model_utils.py

@@ -1,6 +1,6 @@
 import os
 import logging
-from scipy.stats import pearsonr
+from scipy.stats import pearsonr, spearmanr
 import utils  # Sets up useful loggers
 from data import base
 import shutil
@@ -12,7 +12,7 @@ import pandas as pd
 sns.set(style="darkgrid", color_codes=True)
 
 
-class Model():
+class Model:
     def __init__(self, logdir, dataset, config):
         self.training_set = dataset
         if not os.path.isabs(logdir):
@@ -42,14 +42,14 @@ class Model():
         raise NotImplemented
 
     def predict(self, sa, sb):
-        sa = self.preprocessLine(sa)
-        sb = self.preprocessLine(sb)
+        sa = self.preprocess_line(sa)
+        sb = self.preprocess_line(sb)
         return self._predict(sa, sb)
 
-    def preprocessLine(self, sentence):
+    def preprocess_line(self, sentence):
         raise NotImplemented
 
-    def evalScore(self, eval_set=None):
+    def eval_score(self, eval_set=None):
         if eval_set is None:
             eval_set = self.training_set
         self.logger.info("Evaluating on %s", eval_set.name)

### 2017-05-29 src/baseline.py
Data analysis notebook added

@@ -1,7 +1,4 @@
-import pandas as pd
 import nltk
-import numpy as np
-import matplotlib.pyplot as plt
 from data import eval_ds
 from model_utils import Model
 
@@ -16,7 +13,6 @@ class SimpleBaseline(Model):
 
     def preprocessLine(self, sentence):
         return nltk.word_tokenize(sentence)
-
 
 model = SimpleBaseline("log", eval_ds, {})
 model.evalScore()

### 2017-05-29 src/cosine_baseline.py

@@ -1,6 +1,4 @@
-import matplotlib.pyplot as plt
 from sklearn.metrics.pairwise import cosine_similarity
-import pandas as pd
 import nltk
 import numpy as np
 from data import eval_ds, sentence_to_vector

### 2017-05-29 src/data.py

@@ -1,7 +1,3 @@
-import gensim
-import logging
-import utils
-
 """
 If you get this error after importing gensim:
 "Intel MKL FATAL ERROR: Cannot load libmkl_avx2.so or libmkl_def.so"
@@ -9,9 +5,12 @@ try running:
 conda install -f numpy
 from the command line and it magically fixes it.
 """
+import gensim
+import logging
 import numpy as np
 import os
 from dotenv import load_dotenv
+
 
 logger = logging.getLogger(__file__)
 

### 2017-05-29 src/model_utils.py

@@ -5,7 +5,6 @@ import utils  # Sets up useful loggers
 from data import base
 import shutil
 from time import localtime, strftime
-import matplotlib.pyplot as plt
 import seaborn as sns
 import pandas as pd
 
@@ -35,7 +34,7 @@ class Model():
         fh.setFormatter(logging.Formatter(utils.log_fmt))
         self.logger.addHandler(fh)
 
-    def train(self, num_steps):
+    def train(self, train_x, train_y, num_steps):
         raise NotImplemented
 
     def _predict(self, sa, sb):

### 2017-05-30 paper/main.tex
Small update to paper

@@
-\documentclass[12pt]{article}
+% Paper template for TAR 2016
+% (C) 2014 Jan Šnajder, Goran Glavaš, Domagoj Alagić, Mladen Karan
+% TakeLab, FER
+
+\documentclass[10pt, a4paper]{article}
+
+\usepackage{tar2016}
+
+\usepackage[utf8]{inputenc}
+\usepackage[pdftex]{graphicx}
+\usepackage{booktabs}
 \usepackage{amsmath}
+\usepackage{amssymb}
+
+\title{Semantic Textual Similarity using deep learning}
+
+\name{Bruno Gavranovic, Neven Miculinic, Stipan Mikulic}
+
+\address{
+University of Zagreb, Faculty of Electrical Engineering and Computing\\
+Unska 3, 10000 Zagreb, Croatia\\
+\texttt{\{bgavran3,neven.miculinic,?stipan?\}@gmail.com}\\
+}
+
+\abstract{
+In this paper we present our work on Semantic Textual Similarity (STS) problem during Text analysis and retrival (TAR) class at FER.
+STS measure semantic similarity between two sentences.
+Datasets were obtained using human annotators and their average semantic similarity score is given.
+We use both Pearson and Spearman correlation coefficient in our analysis and acchine 1.000, and 1.000 on our test set respectively.
+}
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\maketitleabstract
+
+\section{Introduction}
+
+This section is the introduction to your paper. Introduction should not be too elaborate, as that is what other sections are for (the Introduction should definitely not spill over to the second page).
+
+This is the second paragraph of the introduction.
+In \LaTeX , paragraphs are separated by inserting an empty line in between them.
+Avoid very large paragraphs (larger than half of the page height), but also avoid tiny paragraphs (e.g., one-sentence paragraphs).
+
+\section{Second section}
+
+In scientific papers, this section usually (but not necessarily) briefly describes the related research and what makes the presented approach different from it.
+
+\subsection{First subsection}
+\label{sec:first}
+
+This is a subsection of the second section.
+
+\subsection{Second subsection}
+
+This is the second subsection of the second section. Referencing the (sub)sections in text is performed as follows: ``in Section~\ref{sec:first} we have shown \dots''.
+
+\subsubsection{Sub-subsection example}
+
+This is a sub-subsection. If possible, it is better to avoid sub-subsections.
+
+\section{Extent of the paper}
+
+The paper should have a minimum of 3 and a maximum of 4 pages, plus an additional page for references.
+
+\section{Figures and tables}
+
+\subsection{Figures}
+
+Here is an example on how to include figures in the paper.
+Figures are included in \LaTeX{} code immediately \textit{after} the text in which these figures are referenced.
+Allow \LaTeX{} to place the figure where it believes is best (usually on top of the page of at the position where you would not place the figure).
+Figures are referenced as follows: ``Figure~\ref{fig:figure1} shows \dots''.
+Use tilde (\verb.~.) to prevent separation between the word ``Figure'' and its enumeration.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{drawing.pdf}
+\caption{This is the figure caption.
+Full sentences should be followed with a dot.
+The caption should be placed \textit{below} the figure.
+Caption should be short; details should be explained in the text.}
+\label{fig:figure1}
+\end{center}
+\end{figure}
+
+\subsection{Tables}
+
+There are two types of tables: narrow tables that fit into one column and a wide table that spreads over both columns.
+
+\subsubsection{Narrow tables}
+
+Table~\ref{tab:narrow-table} is an example of a narrow table.
+Do not use vertical lines in tables -- vertical tables have no effect and they make tables visually less attractive.
+We recommend using \textit{booktabs} package for nicer tables.
+
+\begin{table}
+\caption{This is the caption of the table. Table captions should be placed \textit{above} the table.}
+\label{tab:narrow-table}
+\begin{center}
+\begin{tabular}{ll}
+\toprule
+Heading1 & Heading2 \\
+\midrule
+One & First row text \\
+Two   & Second row text \\
+Three   & Third row text \\
+      & Fourth row text \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\subsection{Wide tables}
+
+Table~\ref{tab:wide-table} is an example of a wide table that spreads across both columns. The same can be done for wide figures that should spread across the whole width of the page.
+
+\begin{table*}
+\caption{Wide-table caption}
+\label{tab:wide-table}
+\begin{center}
+\begin{tabular}{llr}
+\toprule
+Heading1 & Heading2 & Heading3\\
+\midrule
+A & A very long text, longer that the width of a single column & $128$\\
+B & A very long text, longer that the width of a single column & $3123$\\
+C & A very long text, longer that the width of a single column & $-32$\\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table*}
+
+\section{Math expressions and formulas}
+
+Math expressions and formulas that appear within the sentence should be written inside the so-called \emph{inline} math environment: $2+3$, $\sqrt{16}$, $h(x)=\mathbf{1}(\theta_1 x_1 + \theta_0>0)$.
+Larger expressions and formulas (e.g., equations) should be written in the so-called \emph{displayed} math environment:
+
+\[
+b^{(i)}_k = \begin{cases}
+1 & \text{if
+    $k = \text{argmin}_j \| \mathbf{x}^{(i)} - \mathbf{\mu}_j \|,$}\\
+0 & \text{otherwise}
+\end{cases}
+\]
+
+Math expressions which you reference in the text should be written inside the \textit{equation} environment:
+
+\begin{equation}\label{eq:kmeans-error}
+J = \sum_{i=1}^N \sum_{k=1}^K
+b^{(i)}_k \| \mathbf{x}^{(i)} - \mathbf{\mu}_k \|^2
+\end{equation}
+
+Now you can reference equation \eqref{eq:kmeans-error}. If the paragraph continues right after the formula
+
+\begin{equation}
+f(x) = x^2 + \varepsilon
+\end{equation}
+
+\noindent like this one does, use the command \emph{noindent} after the equation to remove the indentation of the row.
+
+Multi-letter words in the math environment should be written inside the command \emph{mathit}, otherwise \LaTeX{} will insert spacing between the letters to denote the multiplication of values denoted by symbols.
+For example, compare
+$\mathit{Consistent}(h,\mathcal{D})$ and\\
+$Consistent(h,\mathcal{D})$.
+
+If you need a math symbol, but you don't know the corresponding \LaTeX{} command that generates it, try
+\emph{Detexify}.\footnote{\texttt{http://detexify.kirelabs.org/}}
+
+\section{Referencing literature}
+
+References to other publications should be written in brackets with the last name of the first author and the year of publication, e.g., \citep{chomsky-73}.
+Multiple references are written in sequence, one after another, separated by semicolon and without whitespaces in between, e.g., \citep{chomsky-73,chave-64,feigl-58}.
+References are typically written at the end of the sentence and necessarily before the sentence punctuation.
+
+If the publication is authored by more than one author, only the name of the first author is written, after which abbreviation \emph{et al.}, meaning \emph{et alia}, i.e.,~and others is written as in \citep{johnson-etc}.
+If the publication is authored by only two authors, then the last names of both authors are written \citep{johnson-howells}.
+
+If the name of the author is incorporated into the text of the sentence, it should not be in the brackets (only the year should be there). E.g.,~``\citet{chomsky-73}
+suggested that \dots''. The difference is whether you reference the publication or the author who wrote it.
+
+The list of all literature references is given alphabetically at the end of the paper. The form of the reference depends on the type of the bibliographic unit: conference papers,
+\citep{chave-64}, books \citep{butcher-81}, journal articles
+\citep{howells-51}, doctoral dissertations \citep{croft-78}, and book chapters \citep{feigl-58}.
+
+All of this is automatically produced when using BibTeX. Insert all the BibTeX entries into the file \texttt{tar2016.bib}, and then reference them via their symbolic names.
+
+\section{Conclusion}
+
+Conclusion is the last enumerated section of the paper.
+It should not exceed half of a column and is typically split into 2--3 paragraphs.
+No new information should be presented in the conclusion; this section only summarizes and concludes the paper.
+
+\section*{Acknowledgements}
+
+If suitable, you can include the \textit{Acknowledgements} section before inserting the literature references  in order to thank those who helped you in any way to deliver the paper, but are not co-authors of the paper.
+
+\bibliographystyle{tar2016}
+\bibliography{tar2016}
 

### 2017-05-30 paper/main.tex
Added initial related work

@@
 
-\section{Second section}
+\section{Related work}
 
-In scientific papers, this section usually (but not necessarily) briefly describes the related research and what makes the presented approach different from it.
+STS has short and fruitful history, one with many ideas flying around.
+In its current form it appeared in 2012, on SemVal~\citep{agirre2012semeval} as task 6.
+In 2012, the best system~\citep{bar2012ukp} used lexical similarity and Explicit Semantic Analysis(ESA)~\citep{gabrilovich2007computing}.
+Following year Latent Semantic Analysis(LSA)~\citep{deerwester1990indexing} model~\citep{han2013umbc} with additional external information sources, WordNet and n-gram matching technique.
+
+Following two years~\citep{sultan2014dls} and~\citep{sultan2015dls} dominate the competition with new algorightm -- they align the words between new sentences.
+Other notable approaches come from logic side, it's representative paper being~\citep{beltagy2014probabilistic}.
 

### 2017-05-30 src/char_lstm/lstm.py
Added char_lstm model

@@ -1,0 +1,156 @@
+import tensorflow as tf
+from model_utils import Model
+from sklearn.metrics.pairwise import cosine_similarity
+
+
+class LSTM(Model):
+    max_outputs = 4
+
+    def __init__(self, memory_size, n_layers, out_vector_size, temp, task, model_info):
+        super().__init__(*model_info)
+        self.memory_size = memory_size
+        self.out_vector_size = self.memory_size
+        self.n_layers = n_layers
+        self.out_vector_size = out_vector_size
+        self.temp = temp
+        self.task = task
+
+        one_cell = tf.contrib.rnn.BasicLSTMCell
+        self.lstm_cell = tf.contrib.rnn.MultiRNNCell(
+            [one_cell(self.memory_size, state_is_tuple=False) for _ in range(self.n_layers)])
+
+        self.restore_path = None
+        self.saver = None
+        self.eval_sess = None
+        self.final_state = None
+        self.x = None
+
+    def eval_score_tf(self):
+        with tf.Session() as sess:
+            self.eval_sess = sess
+            self.saver.restore(sess, self.restore_path)
+            self.logger.info("Restored model {} !!!!".format(self.restore_path))
+            graph = sess.graph
+            getopt = graph.get_operation_by_name
+
+            self.final_state = getopt("final_state").outputs[0]
+            self.x = getopt("x").outputs[0]
+            self.eval_score()
+
+    def _predict(self, a, b):
+        sem1 = self.eval_sess.run(self.final_state, feed_dict={self.x: a}).flatten().reshape(1, -1)
+        sem2 = self.eval_sess.run(self.final_state, feed_dict={self.x: b}).flatten().reshape(1, -1)
+        return cosine_similarity(sem1, sem2)[0, 0]
+
+    def preprocess_line(self, sentence):
+        return self.task.text_to_indices(sentence)
+
+    def __call__(self, x, initial_state):
+        """
+        Creates the unrolled RNN specified by cell and transforms it to specified size
+
+        :param x: inputs for all time steps, shape [batch_size, max_time, vector_size]
+        :param initial_state:
+        :return: outputs for all time steps of shape [batch_size, max_time, out_vector_size]
+        """
+        with tf.variable_scope("LSTM"):
+            # sequence_length = tf.stack([sequence_length for _ in range(self.batch_size)])
+            outputs, final_state = tf.nn.dynamic_rnn(self.lstm_cell, x, initial_state=initial_state)
[…]

### 2017-05-30 src/char_lstm/main.py

@@ -1,0 +1,35 @@
+import os
+from src.char_lstm.lstm import LSTM
+from src.char_lstm.proj_path import project_path
+from src.char_lstm.text import Text
+from data import eval_ds
+
+from nltk.corpus import webtext, abc, gutenberg
+
+# file_name = "input.txt"
+# file = Text.open_file(file_name)
+
+# file_list = [webtext.raw("overheard.txt"), abc.raw("rural.txt"), gutenberg.raw("carroll-alice.txt")]
+file = webtext.raw("overheard.txt")
+task = Text(file)
+
+
+class Hp:
+    lstm_memory_size = 512
+    out_vector_size = task.num_chars
+    n_layers = 2
+    temp = 0.7
+
+    batch_size = 128
+    seq_len = 64
+    steps = 10000
+
+
+model_info = ["log_lstm", eval_ds, {"log_overwrite:": True}]
+model = LSTM(Hp.lstm_memory_size, Hp.n_layers, Hp.out_vector_size, Hp.temp, task, model_info)
+
+model.train(Hp, project_path)
+
+# model.restore_path = os.path.join(project_path.base, project_path.log_dir, "May_30__14:22", "train", "model.chpt")
+model.restore_path = project_path.model_path + "-" + str(Hp.steps)
+model.eval_score_tf()

### 2017-05-30 src/char_lstm/text.py

@@ -1,0 +1,52 @@
+import os
+import numpy as np
+from src.char_lstm.proj_path import project_path
+
+
+class Text:
+    def __init__(self, file):
+        self.file = file
+        split_point = int(0.75 * len(self.file))
+        self.file_train, self.file_test = self.file[:split_point], self.file[split_point:]
+
+        self.chars = sorted(list(set(self.file)))
+        self.num_chars = len(self.chars)
+        self.char_to_ind = {c: ind for ind, c in enumerate(self.chars)}
+        self.ind_to_char = {ind: c for c, ind in self.char_to_ind.items()}
+
+        self.file_train_ind = [self.char_to_ind[char] for char in self.file_train]
+        self.file_test_ind = [self.char_to_ind[char] for char in self.file_test]
+
+    def next_batch(self, batch_size, seq_len=20, test=False):
+        if test:
+            source = self.file_test_ind
+        else:
+            source = self.file_train_ind
+        indices = np.random.randint(0, len(source) - seq_len, batch_size)
+        x = np.array([source[ind:ind + seq_len] for ind in indices])
+        return x
+
+    def text_to_indices(self, text):
+        x = np.expand_dims([self.char_to_ind[char] for char in text], axis=0)
+        return x
+
+    def sample_text(self, sess, next_char_tf, next_state_tf, x, initial_state, seed=None, length=20):
+        if seed is None:
+            next_char = np.array([[np.random.randint(0, self.num_chars)]])
+        else:
+            next_char = self.text_to_indices(seed)
+        next_state = None
+        sampled_sentence = self.ind_to_char[next_char[0, 0]]
+        for i in range(length):
+            feed_dict = {x: next_char}
+            if next_state is not None:
+                feed_dict[initial_state] = tuple(next_state)
+            next_char, next_state = sess.run([next_char_tf, next_state_tf], feed_dict=feed_dict)
+            char = self.ind_to_char[next_char[0, 0]]
+            sampled_sentence += char
+        return sampled_sentence, next_state
+
+    @staticmethod
+    def open_file(file_name):
+        text_path = os.path.join(project_path.base, file_name)
+        return open(text_path).read()

### 2017-05-31 src/char_lstm/lstm.py
Fixed bugs, added bigger dataset

@@ -38,8 +38,10 @@ class LSTM(Model):
             self.eval_score()
 
     def _predict(self, a, b):
-        sem1 = self.eval_sess.run(self.final_state, feed_dict={self.x: a}).flatten().reshape(1, -1)
-        sem2 = self.eval_sess.run(self.final_state, feed_dict={self.x: b}).flatten().reshape(1, -1)
+        final_state1 = self.eval_sess.run(self.final_state, feed_dict={self.x: a})
+        final_state2 = self.eval_sess.run(self.final_state, feed_dict={self.x: b})
+        sem1 = final_state1[1].flatten().reshape(1, -1)
+        sem2 = final_state2[1].flatten().reshape(1, -1)
         return cosine_similarity(sem1, sem2)[0, 0]
 
     def preprocess_line(self, sentence):
@@ -104,17 +106,18 @@ class LSTM(Model):
         print("This model has", n_vars, "parameters!")
 
         with tf.Session() as sess:
+            self.eval_sess = sess
             graph = sess.graph
             getopt = graph.get_operation_by_name
             getten = graph.get_tensor_by_name
 
             opt = getopt("optimizer")
             loss = getopt("loss").outputs[0]
-            x = getopt("x").outputs[0]
+            self.x = getopt("x").outputs[0]
             merged = getopt("merged").outputs[0]
             one_char = getopt("one_char").outputs[0]
             # initial_state = getopt("initial_state").outputs[0]
-            final_state = getopt("final_state").outputs[0]
+            self.final_state = getopt("final_state").outputs[0]
 
             if restore_path is not None:
                 self.saver.restore(sess, restore_path)
@@ -128,17 +131,18 @@ class LSTM(Model):
             t = time()
 
             print("Starting...")
+            scores_list = []
             for step in range(hp.steps + 1):
                 data_batch = self.task.next_batch(batch_size=hp.batch_size, seq_len=hp.seq_len)
 
-                _, cost_value = sess.run([opt, loss], feed_dict={x: data_batch})
+                _, cost_value = sess.run([opt, loss], feed_dict={self.x: data_batch})
 
                 if step % 100 == 0:
-                    summary = sess.run(merged, feed_dict={x: data_batch})
+                    summary = sess.run(merged, feed_dict={self.x: data_batch})
                     train_writer.add_summary(summary, step)
 
                     test_data_batch = self.task.next_batch(batch_size=hp.batch_size, seq_len=hp.seq_len, test=True)
[…]

### 2017-05-31 src/char_lstm/main.py

@@ -10,19 +10,20 @@ from nltk.corpus import webtext, abc, gutenberg
 # file = Text.open_file(file_name)
 
 # file_list = [webtext.raw("overheard.txt"), abc.raw("rural.txt"), gutenberg.raw("carroll-alice.txt")]
-file = webtext.raw("overheard.txt")
-task = Text(file)
+file_list = [gutenberg.raw(file) for file in gutenberg.fileids()]
+# file = webtext.raw("overheard.txt")
+task = Text(file_list)
 
 
 class Hp:
     lstm_memory_size = 512
     out_vector_size = task.num_chars
-    n_layers = 2
+    n_layers = 3
     temp = 0.7
 
     batch_size = 128
-    seq_len = 64
-    steps = 10000
+    seq_len = 128
+    steps = 100000
 
 
 model_info = ["log_lstm", eval_ds, {"log_overwrite:": True}]

### 2017-05-31 src/char_lstm/text.py

@@ -2,28 +2,34 @@ import os
 import numpy as np
 from src.char_lstm.proj_path import project_path
 
+def flatten(x): return [x for x in x for x in x]
 
 class Text:
-    def __init__(self, file):
-        self.file = file
-        split_point = int(0.75 * len(self.file))
-        self.file_train, self.file_test = self.file[:split_point], self.file[split_point:]
+    def __init__(self, file_list):
+        self.file_list = file_list
+        self.files_train, self.files_test = [], []
+        for file in self.file_list:
+            split_point = int(0.75 * len(file))
+            self.files_train.append(file[:split_point])
+            self.files_test.append(file[split_point:])
 
-        self.chars = sorted(list(set(self.file)))
+        self.chars = sorted(list(set(flatten(self.file_list))))
         self.num_chars = len(self.chars)
         self.char_to_ind = {c: ind for ind, c in enumerate(self.chars)}
         self.ind_to_char = {ind: c for c, ind in self.char_to_ind.items()}
 
-        self.file_train_ind = [self.char_to_ind[char] for char in self.file_train]
-        self.file_test_ind = [self.char_to_ind[char] for char in self.file_test]
+        self.file_train_ind = [[self.char_to_ind[char] for char in file] for file in self.files_train]
+        self.file_test_ind = [[self.char_to_ind[char] for char in file] for file in self.files_test]
 
     def next_batch(self, batch_size, seq_len=20, test=False):
         if test:
             source = self.file_test_ind
         else:
             source = self.file_train_ind
-        indices = np.random.randint(0, len(source) - seq_len, batch_size)
-        x = np.array([source[ind:ind + seq_len] for ind in indices])
+        file_ind = np.random.randint(0, len(source))
+        file = source[file_ind]
+        indices = np.random.randint(0, len(file) - seq_len, batch_size)
+        x = np.array([file[ind:ind + seq_len] for ind in indices])
         return x
 
     def text_to_indices(self, text):

### 2017-05-31 paper/main.tex
Small changes to latex paper

@@
 
+\graphicspath{{img/}}
+
 \title{Semantic Textual Similarity using deep learning}
@@
 Unska 3, 10000 Zagreb, Croatia\\
-\texttt{\{bgavran3,neven.miculinic,?stipan?\}@gmail.com}\\
+\texttt{\{bruno.gavranovic,neven.miculinic,stipan.mikulic\}@fer.hr}\\
 }
@@
 In this paper we present our work on Semantic Textual Similarity (STS) problem during Text analysis and retrival (TAR) class at FER. 
-STS measure semantic similarity between two sentences.
+STS measures semantic similarity between two sentences. 
+We use two approaches: an SVM model with feature preprocessing and a deep learning model trained on a language modelling task.
 Datasets were obtained using human annotators and their average semantic similarity score is given. 
-We use both Pearson and Spearman correlation coefficient in our analysis and acchine 1.000, and 1.000 on our test set respectively.
+We use both Pearson and Spearman correlation coefficient in our analysis and achieve 1.000, and 1.000 on our test set respectively.
 }
@@
 \section{Introduction}
+
+STS has many applications in NLP bla bla bla
 
@@
 
-Following two years~\citep{sultan2014dls} and~\citep{sultan2015dls} dominate the competition with new algorightm -- they align the words between new sentences.
-Other notable approaches come from logic side, it's representative paper being~\citep{beltagy2014probabilistic}.
-
-\subsection{First subsection}
-\label{sec:first}
-
-This is a subsection of the second section.
-
-\subsection{Second subsection}
-
-This is the second subsection of the second section. Referencing the (sub)sections in text is performed as follows: ``in Section~\ref{sec:first} we have shown \dots''.
-
-\subsubsection{Sub-subsection example}
-
-This is a sub-subsection. If possible, it is better to avoid sub-subsections.
+Following two years~\citep{sultan2014dls} and~\citep{sultan2015dls} dominate the competition with new algorithm -- they align the words between new sentences.
+Other notable approaches come from logic side, its representative paper being~\citep{beltagy2014probabilistic}.
 
@@
 
-The paper should have a minimum of 3 and a maximum of 4 pages, plus an additional page for references.
+Our contribution consists of implementation of two machine learning models: SVM model and a deep learning model.
 
-\section{Figures and tables}
+\subsection{SVM}
 
-\subsection{Figures}
-
-Here is an example on how to include figures in the paper.
-Figures are included in \LaTeX{} code immediately \textit{after} the text in which these figures are referenced.
-Allow \LaTeX{} to place the figure where it believes is best (usually on top of the page of at the position where you would not place the figure).
-Figures are referenced as follows: ``Figure~\ref{fig:figure1} shows \dots''.
-Use tilde (\verb.~.) to prevent separation between the word ``Figure'' and its enumeration.
+\subsection{Deep learning}
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{drawing.pdf}
-\caption{This is the figure caption.
-Full sentences should be followed with a dot.
-The caption should be placed \textit{below} the figure.
-Caption should be short; details should be explained in the text.}
-\label{fig:figure1}
+\includegraphics[width=\columnwidth]{only_2nd_layer.png}
+\caption{Depiction of blabla}
+\label{fig:lstm_2nd_layer}
 \end{center}

### 2017-05-31 src/char_lstm/lstm.py

@@ -40,8 +40,8 @@ class LSTM(Model):
     def _predict(self, a, b):
         final_state1 = self.eval_sess.run(self.final_state, feed_dict={self.x: a})
         final_state2 = self.eval_sess.run(self.final_state, feed_dict={self.x: b})
-        sem1 = final_state1[1].flatten().reshape(1, -1)
-        sem2 = final_state2[1].flatten().reshape(1, -1)
+        sem1 = final_state1[-1].flatten().reshape(1, -1)
+        sem2 = final_state2[-1].flatten().reshape(1, -1)
         return cosine_similarity(sem1, sem2)[0, 0]
 
     def preprocess_line(self, sentence):

### 2017-05-31 src/char_lstm/main.py

@@ -1,7 +1,7 @@
 import os
 from src.char_lstm.lstm import LSTM
 from src.char_lstm.proj_path import project_path
-from src.char_lstm.text import Text
+from src.char_lstm.onehot import OneHot
 from data import eval_ds
 
 from nltk.corpus import webtext, abc, gutenberg
@@ -12,7 +12,7 @@ from nltk.corpus import webtext, abc, gutenberg
 # file_list = [webtext.raw("overheard.txt"), abc.raw("rural.txt"), gutenberg.raw("carroll-alice.txt")]
 file_list = [gutenberg.raw(file) for file in gutenberg.fileids()]
 # file = webtext.raw("overheard.txt")
-task = Text(file_list)
+task = OneHot(file_list)
 
 
 class Hp:

### 2017-05-31 src/char_lstm/onehot.py

@@ -2,9 +2,11 @@ import os
 import numpy as np
 from src.char_lstm.proj_path import project_path
 
+
 def flatten(x): return [x for x in x for x in x]
 
-class Text:
+
+class OneHot:
     def __init__(self, file_list):
         self.file_list = file_list
         self.files_train, self.files_test = [], []

### 2017-06-01 paper/main.tex
Added initial Introduction

@@
 
-STS has many applications in NLP bla bla bla
+\begin{verbatim}
+Two dogs play in the grass.
+Two dogs playing in the snow.
+\end{verbatim}
 
-This section is the introduction to your paper. Introduction should not be too elaborate, as that is what other sections are for (the Introduction should definitely not spill over to the second page).
+At the heart of STS, semantic textual similarity, is assigning numerical score on two text similarity.
+By text, it could mean full document, paragraph, or in this paper's case only one sentence.
+Beginning quote showcases one such sentance pairs, where human annotators rated 2.8 semantically similar on 0-5 scale.
+Full score semantics you can read in table~\ref{tab:sts-score}
 
-This is the second paragraph of the introduction.
-In \LaTeX , paragraphs are separated by inserting an empty line in between them.
-Avoid very large paragraphs (larger than half of the page height), but also avoid tiny paragraphs (e.g., one-sentence paragraphs).
+\begin{table*}
+\caption{STS score description, adapted from~\citep{agirre2016semeval}}
+\label{tab:sts-score}
+\begin{center}
+\begin{tabular}{llr}
+\toprule
+Score & Explanation\\
+\midrule
+5 & \textit{Two sentences are completely equivalent}\\
+& The bird is bathing in the sink\\
+& Birdie is washing itself in the water basin.\\
+\midrule
+4 & \textit{The two sentences are mostly equivalent, but some unimportant details differ}\\
+& In May 2010, the troops attempted to invade
+Kabul. \\
+& The US army invaded Kabul on May 7th last
+year, 2010.\\
+\midrule
+3 & \textit{The two sentences are roughly equivalent, but some important information differs/missing.}\\
+& John said he is considered a witness but not a
+suspect.\\
+& ``He is not a suspect anymore.'' John said.\\
+\midrule
+2 & \textit{The two sentences are not equivalent, but share some details.} \\
+&They flew out of the nest in groups. \\
+&They flew into the nest together. \\
+\midrule
+1 & \textit{The two sentences are not equivalent, but are on the same topic.}\\
+& The woman is playing the violin.\\
+& The young lady enjoys listening to the guitar.\\
+\midrule
+0 & \textit{The two sentences are completely dissimilar.}\\
+& John went horse back riding at dawn with a whole group of friends.\\
+& Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.\\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table*}
 

### 2017-06-01 paper/main.tex
Small introduction update
TODO:

See http://www.aclweb.org/anthology/S12-1060 and make it more reference
filled and complete

@@
 By text, it could mean full document, paragraph, or in this paper's case only one sentence.
-Beginning quote showcases one such sentance pairs, where human annotators rated 2.8 semantically similar on 0-5 scale.
+Beginning quote showcases one such sentance pairs, where human annotators rated 2.8 semantically similar on 0-5 scale, where 0 mean full semantic dissimilarity, and 5 full equivalence with various shades in between.
 Full score semantics you can read in table~\ref{tab:sts-score}
+
+STS developed system and techniqus have many further applications, transfer learning of sorts.
+Machine translation(MT), Summarization, Generation and Question Answering(QA) are some of them.
+Often new techniques invented in STS context generlize to earlier mentioned domains, as well as NLP fiels as a whole.
 

### 2017-06-02 src/data.py
Svm model done

@@ -20,9 +20,9 @@ load_dotenv(os.environ.get(
     os.path.join(base, 'src', 'local.env')
 ))
 
-w2v_path = os.path.join(base, os.environ['W2V_PATH'])
-eval_data_in_path = os.path.join(base, os.environ["EVAL_IN"])
-eval_data_out_path = os.path.join(base, os.environ["EVAL_OUT"])
+w2v_path = os.path.join(base, os.environ.get('W2V_PATH', ""))
+eval_data_in_path = os.path.join(base, os.environ.get("EVAL_IN", ""))
+eval_data_out_path = os.path.join(base, os.environ.get("EVAL_OUT", ""))
 
 
 class DataSet:

### 2017-06-02 src/svm.py

@@ -1,0 +1,297 @@
+import nltk
+import numpy as np
+import re
+import os
+import seaborn as sns
+import pandas as pd
+from nltk.corpus.reader import WordNetError
+from sklearn.metrics.pairwise import cosine_similarity
+from sklearn.model_selection import GridSearchCV, train_test_split
+from sklearn.preprocessing import StandardScaler
+from sklearn.svm import SVR
+from nltk.stem import WordNetLemmatizer
+from nltk import FreqDist
+from nltk.corpus import brown, wordnet
+from scipy.stats import pearsonr
+
+
+wordnet_lemmatizer = WordNetLemmatizer()
+stopwords = set(nltk.corpus.stopwords.words("english"))
+
+tagger = nltk.tag.pos_tag
+frequency_list = FreqDist(i.lower() for i in brown.words())
+all_words_count = 0
+for i in frequency_list:
+    all_words_count += frequency_list[i]
+
+
+def get_words(sentence):
+    return [i.strip('., ') for i in sentence.split(' ')]
+
+
+with open('word_to_vec', 'r') as f:
+    embeddings = {}
+    for line in f.readlines():
+        args = get_words(line.strip("\n\t "))
+        embeddings[args[0]] = [float(i) for i in args[1:]]
+
+
+def get_ngram(word, n):
+    ngrams = []
+    word_len = len(word)
+    for i in range(word_len - n + 1):
+        ngrams.append(word[i: i + n])
+    return ngrams
+
+
+def get_lists_intersection(s1, s2):
+    s1_s2 = []
+    for i in s1:
+        if i in s2:
+            s1_s2.append(i)
+    return s1_s2
+
+
+def overlap(sentence1_ngrams, sentence2_ngrams):
+    s1_len = len(sentence1_ngrams)
+    s2_len = len(sentence2_ngrams)
+    if s1_len == 0 and s2_len == 0:
+        return 0
+    s1_s2_len = max(1, len(get_lists_intersection(sentence2_ngrams, sentence1_ngrams)))
+    return 2 / (s1_len / s1_s2_len + s2_len / s1_s2_len)
+
+
+def get_ngram_feature(sentence1, sentence2, n):
+    sentence1_ngrams = []
+    sentence2_ngrams = []
+
+    for word in sentence1:
+        sentence1_ngrams.extend(get_ngram(word, n))
+
+    for word in sentence2:
[…]

### 2017-06-03 paper/main.tex

@@
 \usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 
+\subsection{Data analysis}
+We used dataset from SemEval 2017 
+Task 1\footnote{http://alt.qcri.org/semeval2017/task1/}. 
+The dataset has 250 instances.
+Each instance consists of two sentences and our tash is to predict semantic similarity between them.
+In order to get some deeper insights on data we did some minor data analysis.
+We got to conclusion that we are dealing with very short text.
+Also, we concluded that the smaller difference between lengths of sentences is, the more similar they are.
+This can be seen clearer on following plots.
+
+\begin{table}
+\caption{Stats about lengths of sentences}
+\label{tab:narrow-table}
+\begin{center}
+\begin{tabular}{cll}
+\toprule
+& Without stopwords & With stopwords \\
+\midrule
+Sentence length & 33.004 & 43.838 \\
+Tokens in sentence & 5.488 & 8.702 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\columnwidth]{sen_len_no_stop.png}
+\caption{Sentence lengths without stopwords}
+\label{fig:lstm_2nd_layer}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\columnwidth]{sen_len_with_stop.png}
+\caption{Sentence lengths with stopwords}
+\label{fig:lstm_2nd_layer}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\columnwidth]{tokens_no_stop.png}
+\caption{Number of tokens without stopwords}
+\label{fig:lstm_2nd_layer}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\columnwidth]{tokens_with_stop.png}
+\caption{Number of tokens with stopwords}
+\label{fig:lstm_2nd_layer}
+\end{center}
+\end{figure}
+
+\newpage
 \subsection{SVM}
+
+\subsubsection{Features}
+All of the features we used are from \citep{Saric2012TakeLabSF} paper.
+\paragraph{Ngram Overlap \\}
+Let S1 and S2 be the sets of consecutive ngrams in the first and the second sentence, respectively. The ngram overlap is computed for unigrams, bigrams, and trigrams. It is defined as follows:
+\begin{equation}\label{eq:ngo}
+ngo(S_1, S_2) = 2 \cdot \bigg( \frac{|S_1|}{|S_1 \cap S_2|}+\frac{|S_2|}{|S_1 \cap S_2|}\bigg)^{-1}
+\end{equation}
+
+The ngram overlap is the harmonic mean of the degree
+to which the second sentence covers the first
+and vice versa.\citep{Saric2012TakeLabSF}
+
+\paragraph{WordNet-Augmented Word Overlap \\}
+In order to determine some semantical meaning from words we define the WordNet augmented coverage $ PWN(\cdot, \cdot) $:
+\begin{equation}\label{eq:pwn}
+P_{WN}(S_1, S_2) = \frac{1}{|S_2|} \sum_{w \in S_1} score(w, S_2)
+\end{equation}
+
+\begin{equation}\label{eq:pwn-score}
+score(w, S) = \begin{cases}
+1 & \text{if $w \in S $}\\
+\max\limits_{w^{\rq} \in S} sim(w, w^{\rq}) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+where $sim(\cdot, \cdot)$ represents the WordNet path length
+similarity. The WordNet-augmented word overlap
+feature is defined as a harmonic mean of
+$PWN(S_1, S_2)$ and $PWN(S_2, S_1)$.\citep{Saric2012TakeLabSF}
+
+\paragraph{Weighted Word Overlap \\}
+We define information contente measure to give some words more importance:
+\begin{equation}\label{eq:inf-content}
+ic(w) = ln \frac{\sum_{w^{\rq} \in C} freq(w^{\rq}) }{freq(w)}
+\end{equation}
+where C is the set of words in the corpus and
+freq(w) is the frequency of the word w in the corpus. Since we got very small dataset we used \emph{nltk brown} dataset to calculate word frequency distribution.
+The weighted word coverage of the second sentence by the first sentence is given by:
+\begin{equation}\label{eq:wwc}
+wwc(S_1, S_2) = \frac{\sum_{w \in S_1 \cap S_2} ic(w)}{\sum_{w^{\rq} \in S_2} ic(w^{\rq})}
+\end{equation}
+where $S_1$ and $S_2$ are words in sentences.\\
+The \emph{weighted word overlap} between two sentences
+is calculated as the harmonic mean of the
+$wwc(S_1, S_2)$ and $wwc(S_2, S_1)$.\citep{Saric2012TakeLabSF}
+
+\paragraph{Number of tokens difference \\}
+Difference between number of words per sentence is defined as:
+\begin{equation}\label{eq:word-diff}
+diff(S_1, S_2) = abs(|S_1| - |S_2|)
+\end{equation}
+where $S_1$ and $S_2$ are words in sentences.
+\paragraph{Vector Space Sentence Similarity \\}
+We define each sentence as vector $u(\cdot)$ by summing all word embeddings in the sentence S: $ u(S) = \sum_{w \in S} x_w$ where $x_w$ is word embedding for each word. Another similar
+representation $u_w(\cdot)$ uses the information content
+$ic(w)$ to weigh the word embedding vector of each word
+before summation: $u_w(S) = \sum_{w \in S} ic(w) \cdot x_w.$
+We use $|cos(u(S_1), u(S_2))|$ and $|cos(u_w(S_1), u_w(S_2))|$ for the vector space sentence similarity features. \citep{Saric2012TakeLabSF}
+\paragraph{Shallow NER Features \\}
+For this feature we simply count all words that are capitalized.
+\paragraph{Numbers Overlap \\}
+For numbers overlap we define following three feature: $log(1+|N_1|+|N_2|)$, $2\cdot|N_1 \cap N_2|/(|N_1|+|N_2|)$ and $N_1 \subseteq N_2 \vee N_2 \subseteq N_1$
+where $N_1$ and $N_2$ are sets of numbers in two sentences. We treat all numbers as decimal numbers.
+
+\subsubsection{Model}
+All features were preprocessed with StandardScaler\footnote{http://scikit-learn.org/stable/modules/generated/sklearn.preprocessing.StandardScaler.html}. 
+\subsubsection{Evaluation}
 
@@
 
+\newpage
 \begin{figure}
@@
 \bibliography{tar2016}
+\nocite{*}
 

### 2017-06-04 paper/main.tex
Svm model done

@@
 \usepackage{amssymb}
+\usepackage{float}
 \usepackage{graphicx}
@@
 \end{verbatim}
-
 At the heart of STS, semantic textual similarity, is assigning numerical score on two text similarity.
@@
 Full score semantics you can read in table~\ref{tab:sts-score}
-
 STS developed system and techniqus have many further applications, transfer learning of sorts.
@@
 
-\begin{figure}[h]
-\begin{center}
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[t]{0.5\textwidth}
         \includegraphics[width=\columnwidth]{sen_len_no_stop.png}
 		\caption{Sentence lengths without stopwords}
-\label{fig:lstm_2nd_layer}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
+    \end{subfigure}%
+    ~ 
+    \begin{subfigure}[t]{0.5\textwidth}
         \includegraphics[width=\columnwidth]{sen_len_with_stop.png}
 		\caption{Sentence lengths with stopwords}
-\label{fig:lstm_2nd_layer}
-\end{center}
-\end{figure}
+    \end{subfigure}
+    ~ 
+    \caption{Sentence length stats}
+\end{figure*}
 
-\begin{figure}[h]
-\begin{center}
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[t]{0.47\textwidth}
 		\includegraphics[width=\columnwidth]{tokens_no_stop.png}
 		\caption{Number of tokens without stopwords}
-\label{fig:lstm_2nd_layer}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[t]{0.49\textwidth}
 		\includegraphics[width=\columnwidth]{tokens_with_stop.png}
 		\caption{Number of tokens with stopwords}
-\label{fig:lstm_2nd_layer}
-\end{center}
-\end{figure}
+    \end{subfigure}
+    ~ 
+    \caption{Number of tokens stats}
+\end{figure*}
+
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=\columnwidth]{sen_len_no_stop.png}
+%\caption{Sentence lengths without stopwords}
+%\label{fig:lstm_2nd_layer}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=\columnwidth]{sen_len_with_stop.png}
+%\caption{Sentence lengths with stopwords}
+%\label{fig:lstm_2nd_layer}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=\columnwidth]{tokens_no_stop.png}
+%\caption{Number of tokens without stopwords}
+%\label{fig:lstm_2nd_layer}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=\columnwidth]{tokens_with_stop.png}
+%\caption{Number of tokens with stopwords}
+%\label{fig:lstm_2nd_layer}
+%\end{center}
+%\end{figure}
 
@@
 All features were preprocessed with StandardScaler\footnote{http://scikit-learn.org/stable/modules/generated/sklearn.preprocessing.StandardScaler.html}.
+For selection of best model we used nested cross validation.
+Both inner and outer CV used 5-Fold cross validation.
+We optimized following parameters of SVR model:
+
+\[ kernel = \{ linear, rbf \} \]
+\[ C = \{ 2^{-7},2^{-6}, ..., 2^6 \} \]
+\[ gamma = \{ 2^{-5}, 2^{-4},..., 2^2 \} \]
+
 \subsubsection{Evaluation}
+Since we used nested cross validation evaluation metrics are averaged over all testing folds.
+We used R2\footnote{http://scikit-learn.org/stable/modules/generated/sklearn.metrics.r2\_score.html} and Pearson correlation
+\footnote{https://docs.scipy.org/doc/scipy-0.14.0/reference/generated/scipy.stats.pearsonr.html} as evaluation metrics. Nested CV scores are dispalyed on following plots:
+
+\begin{figure*}
+\begin{center}
+	\centering
+	\includegraphics[scale=0.36]{R2.png}
+	\caption{R2 score}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+	\centering
+	\includegraphics[scale=0.36]{Pearson.png}
+	\caption{Pearson correlation score}
+\end{center}
+\end{figure*}
+
+\begin{table}
+\caption{Averaged scores}
+\label{tab:narrow-table}
+\begin{center}
+\begin{tabular}{cll}
+\toprule
+& R2 & Pearson correlation\\
+\midrule
+score &  & 0.7795 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\subsection{Tables}
-
-There are two types of tables: narrow tables that fit into one column and a wide table that spreads over both columns.
-
-\subsubsection{Narrow tables}
-
-Table~\ref{tab:narrow-table} is an example of a narrow table.
-Do not use vertical lines in tables -- vertical tables have no effect and they make tables visually less attractive.
-We recommend using \textit{booktabs} package for nicer tables.
-
-\begin{table}
-\caption{This is the caption of the table. Table captions should be placed \textit{above} the table.}
-\label{tab:narrow-table}
-\begin{center}
-\begin{tabular}{ll}
-\toprule
-Heading1 & Heading2 \\
-\midrule
-One & First row text \\
-Two   & Second row text \\
-Three   & Third row text \\
-      & Fourth row text \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Wide tables}
-
-Table~\ref{tab:wide-table} is an example of a wide table that spreads across both columns. The same can be done for wide figures that should spread across the whole width of the page.
-
-\begin{table*}
-\caption{Wide-table caption}
-\label{tab:wide-table}
-\begin{center}
-\begin{tabular}{llr}
-\toprule
-Heading1 & Heading2 & Heading3\\
-\midrule
-A & A very long text, longer that the width of a single column & $128$\\
-B & A very long text, longer that the width of a single column & $3123$\\
-C & A very long text, longer that the width of a single column & $-32$\\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table*}
-
-\section{Math expressions and formulas}
-
-Math expressions and formulas that appear within the sentence should be written inside the so-called \emph{inline} math environment: $2+3$, $\sqrt{16}$, $h(x)=\mathbf{1}(\theta_1 x_1 + \theta_0>0)$.
-Larger expressions and formulas (e.g., equations) should be written in the so-called \emph{displayed} math environment:
-
-\[
-b^{(i)}_k = \begin{cases}
-1 & \text{if
-    $k = \text{argmin}_j \| \mathbf{x}^{(i)} - \mathbf{\mu}_j \|,$}\\
-0 & \text{otherwise}
-\end{cases}
-\]
-
-Math expressions which you reference in the text should be written inside the \textit{equation} environment:
-
-\begin{equation}\label{eq:kmeans-error}
-J = \sum_{i=1}^N \sum_{k=1}^K
-b^{(i)}_k \| \mathbf{x}^{(i)} - \mathbf{\mu}_k \|^2
-\end{equation}
-
-Now you can reference equation \eqref{eq:kmeans-error}. If the paragraph continues right after the formula
-
-\begin{equation}
-f(x) = x^2 + \varepsilon
-\end{equation}
-
-\noindent like this one does, use the command \emph{noindent} after the equation to remove the indentation of the row.
-
-Multi-letter words in the math environment should be written inside the command \emph{mathit}, otherwise \LaTeX{} will insert spacing between the letters to denote the multiplication of values denoted by symbols.
-For example, compare
-$\mathit{Consistent}(h,\mathcal{D})$ and\\
-$Consistent(h,\mathcal{D})$.
-
-If you need a math symbol, but you don't know the corresponding \LaTeX{} command that generates it, try
-\emph{Detexify}.\footnote{\texttt{http://detexify.kirelabs.org/}}
-
-\section{Referencing literature}
-
-References to other publications should be written in brackets with the last name of the first author and the year of publication, e.g., \citep{chomsky-73}.
-Multiple references are written in sequence, one after another, separated by semicolon and without whitespaces in between, e.g., \citep{chomsky-73,chave-64,feigl-58}.
-References are typically written at the end of the sentence and necessarily before the sentence punctuation.
-
-If the publication is authored by more than one author, only the name of the first author is written, after which abbreviation \emph{et al.}, meaning \emph{et alia}, i.e.,~and others is written as in \citep{johnson-etc}.
-If the publication is authored by only two authors, then the last names of both authors are written \citep{johnson-howells}.
-
-If the name of the author is incorporated into the text of the sentence, it should not be in the brackets (only the year should be there). E.g.,~``\citet{chomsky-73}
-suggested that \dots''. The difference is whether you reference the publication or the author who wrote it.
-
-The list of all literature references is given alphabetically at the end of the paper. The form of the reference depends on the type of the bibliographic unit: conference papers,
-\citep{chave-64}, books \citep{butcher-81}, journal articles
-\citep{howells-51}, doctoral dissertations \citep{croft-78}, and book chapters \citep{feigl-58}.
-
-All of this is automatically produced when using BibTeX. Insert all the BibTeX entries into the file \texttt{tar2016.bib}, and then reference them via their symbolic names.
-
 \section{Conclusion}
-
-Conclusion is the last enumerated section of the paper.
-It should not exceed half of a column and is typically split into 2--3 paragraphs.
-No new information should be presented in the conclusion; this section only summarizes and concludes the paper.
-
-\section*{Acknowledgements}
-
-If suitable, you can include the \textit{Acknowledgements} section before inserting the literature references  in order to thank those who helped you in any way to deliver the paper, but are not co-authors of the paper.
 

### 2017-06-04 src/svm.py

@@ -5,14 +5,16 @@ import os
 import seaborn as sns
 import pandas as pd
 from nltk.corpus.reader import WordNetError
+from sklearn.metrics import r2_score, make_scorer
 from sklearn.metrics.pairwise import cosine_similarity
-from sklearn.model_selection import GridSearchCV, train_test_split
+from sklearn.model_selection import GridSearchCV, train_test_split, cross_val_score, KFold
 from sklearn.preprocessing import StandardScaler
 from sklearn.svm import SVR
 from nltk.stem import WordNetLemmatizer
 from nltk import FreqDist
 from nltk.corpus import brown, wordnet
 from scipy.stats import pearsonr, spearmanr
+from matplotlib import pyplot as plt
 
 
 wordnet_lemmatizer = WordNetLemmatizer()
@@ -97,7 +99,7 @@ def get_numbers_feature(sentence1, sentence2):
 def get_shallow_features(sentence):
     counter = 0
     for word in sentence:
-        if len(word) > 1 and re.match("[A-Z].*]", word):
+        if len(word) > 1 and (re.match("[A-Z].*]", word) or re.match("\.[A-Z]+]", word)):
             counter += 1
     return counter
 
@@ -146,7 +148,7 @@ def harmonic_mean(s1, s2):
 
 
 def get_wordnet_pos(treebank_tag):
-    if treebank_tag.startswith('V'):
+    if treebank_tag.startswith('A') or treebank_tag.startswith('JJ'):
         return wordnet.ADJ
     elif treebank_tag.startswith('V'):
         return wordnet.VERB
@@ -160,7 +162,7 @@ def get_wordnet_pos(treebank_tag):
 
 def get_synset(word):
     try:
-        return wordnet.synset(word + "." + get_wordnet_pos(tagger(word)[0][1]) + ".01")
+        return wordnet.synset(word + "." + get_wordnet_pos(tagger([word])[0][1]) + ".01")
     except (WordNetError, IndexError):
         return 0
 
@@ -218,6 +220,75 @@ def feature_vector(a, b):
     return fvec
 
 
+def pearson(y_true, y_predicted):
+    return pearsonr(y_true, y_predicted)[0]
+
+
+def nested_cv(name, scorer=None):
+    train_x = []
+    train_y = []
+    with open('../data/train-en-en.in', 'r') as f:
+        for line in f.readlines():
+            train_x.append([get_words(i.strip('\n\t ')) for i in line.split('\t')])
+
+    with open('../data/train-en-en.out', 'r') as f:
[…]

### 2017-06-04 paper/main.tex
latex updated
Small fixes

@@
 \subsubsection{Model}
-All features were preprocessed with StandardScaler\footnote{http://scikit-learn.org/stable/modules/generated/sklearn.preprocessing.StandardScaler.html}.
-For selection of best model we used nested cross validation.
-Both inner and outer CV used 5-Fold cross validation.
+All features were preprocessed to zero mean and unit variance.
+For selection of best model we used nested cross validation, with Grid search in inner folds.
+Both inner and outer CV used 5-Fold.
 We optimized following parameters of SVR model:
@@
 \begin{center}
-\begin{tabular}{cll}
+\begin{tabular}{ccc}
 \toprule
-& R2 & Pearson correlation\\
+& R2 & Pearson\\
 \midrule
-score &  & 0.7795 \\
+score & 0.5898 & 0.7795 \\
 \bottomrule

### 2017-06-04 paper/main.tex
Added conclusion

@@
 Datasets were obtained using human annotators and their average semantic similarity score is given.
-We use both Pearson and Spearman correlation coefficient in our analysis and achieve 1.000, and 1.000 on our test set respectively.
+We use both Pearson and Spearman correlation coefficient in our analysis and achieve 0.7795, and <TODO> on our test set respectively.
 }
@@
 Since we used nested cross validation evaluation metrics are averaged over all testing folds.
-We used R2\footnote{http://scikit-learn.org/stable/modules/generated/sklearn.metrics.r2\_score.html} and Pearson correlation
-\footnote{https://docs.scipy.org/doc/scipy-0.14.0/reference/generated/scipy.stats.pearsonr.html} as evaluation metrics. Nested CV scores are dispalyed on following plots:
+We used R2 and Pearson correlation as evaluation metrics.
+Nested CV scores are dispalyed on following plots:
 
@@
 
+As presented in this paper, we tried two way approach, both SVM and deep learning based one.
+As we're only students, dabbling with NLP giants, our results aren't considerably behind other papers and SemVal winner.
+Further improvement we see in bigger dataset, combining approaches with model ensables, and even combining various features with CRF, a techinque known for its good performance and resilience to linearly correlated features.
+
+Furthermore additional improvement could be made in bigger model search, both hyperparameters and basic model (e.g.
+Decision trees regression, logistic regression, etc.) using powerful Bayesian hyperoptimization frameworks such as Hyperopt~\citep{bergstra2013hyperopt}
+
 \bibliographystyle{tar2016}

### 2017-06-04 paper/main.tex
Removed spearman coefficient from abstract..because we don't have it
Minor tweaks

@@
 \usepackage{subcaption}
+\usepackage{tabularx}
 
@@
 Datasets were obtained using human annotators and their average semantic similarity score is given.
-We use both Pearson and Spearman correlation coefficient in our analysis and achieve 0.7795, and <TODO> on our test set respectively.
+We use Pearson correlation coefficient in our analysis and achieve 0.7795 on our test set respectively.
 }
@@
 
-\begin{table*}
+\begin{table}[h!]
 \caption{STS score description, adapted from~\citep{agirre2016semeval}}
@@
 \begin{center}
-\begin{tabular}{llr}
+\begin{tabularx}{0.45\textwidth}{llr}
 \toprule
@@
 \midrule
-4 & \textit{The two sentences are mostly equivalent, but some unimportant details differ}\\
-& In May 2010, the troops attempted to invade
-Kabul. \\
-& The US army invaded Kabul on May 7th last
-year, 2010.\\
+4 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
+  & \multicolumn{1}{m{6cm}}{In May 2010, the troops attempted to invade
+Kabul.} \\
+& \multicolumn{1}{m{6cm}}{The US army invaded Kabul on May 7th last
+year, 2010.} \\
 \midrule
-3 & \textit{The two sentences are roughly equivalent, but some important information differs/missing.}\\
-& John said he is considered a witness but not a
-suspect.\\
+3 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
+  & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
+suspect.}\\
 & ``He is not a suspect anymore.'' John said.\\
 \midrule
-2 & \textit{The two sentences are not equivalent, but share some details.} \\
+2 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
 &They flew out of the nest in groups. \\
@@
 \midrule
-1 & \textit{The two sentences are not equivalent, but are on the same topic.}\\
+1 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
 & The woman is playing the violin.\\
@@
 0 & \textit{The two sentences are completely dissimilar.}\\
-& John went horse back riding at dawn with a whole group of friends.\\
-& Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.\\
+  & \multicolumn{1}{m{6cm}}{John went horse back riding at dawn with a whole group of friends.}\\
+  & \multicolumn{1}{m{6cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
 \bottomrule
-\end{tabular}
+\end{tabularx}
 \end{center}
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[h!]
     \centering
@@
 
-\begin{figure*}
+\begin{figure*}[h!]
 \begin{center}
@@
 \subsection{Deep learning}
+
+Deep neural networks have been shown to perform well on language modelling tasks
+(insert citation).
+Use of highly non-linear functions in the models allows for learning of information on a high level of abstraction.
+Semantic textual similarity bla bla idem spavat sutra ovo dovrsim.
+
+\subsection{All results}
+
+Table \ref{tab:all_results} shows the performance of all tested models.
+Bla bla we draw some conclusion from this?
+\begin{table}
+	\caption{Model comparison}
+\label{tab:all_results}
+\begin{center}
+\begin{tabular}{ccc}
+\toprule
+Model & Pearson coefficient \\
+\midrule
+Jaccard similarity & 0.73 \\
+Cosine baseline & 0.73 \\
+LSTM & 0.59 \\
+SVM & 0.78 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 

### 2017-06-05 paper/main.tex
Added baseline descriptions

@@
 
-\name{Bruno Gavranovic, Neven Miculinic, Stipan Mikulic}
+\name{Bruno Gavranović, Neven Miculinić, Stipan Mikulić}
 
@@
 %\end{figure}
+\subsection{Baseline models}
 
-\newpage
+To get an idea of how well our models perform, we used two baseline models.
+Neither of them is a machine learning model; in other words, they do not learn from data, but are rather simple statistics measures.
+
+\subsubsection{Jaccard similarity baseline}
+
+One of the simplest approaches to measuring semantic similarity is the Jaccard rindex.
+It is a statistic used for comparing similarity and diversity of finite sets.
+Let $J(S_1, S_2)$ represent the Jaccard index between two sentences $S_1$ and $S_2$.
+The similariy between sentences is defined as:
+\begin{equation}
+	\label{eq:jaccard}
+	J(S_1, S_2) = \frac{|S_1 \cap S_2|}{|S_1 \cup S_2|},
+\end{equation}
+
+where the elements of $S_1$ and $S_2$ are their words, in lowercase.
+
+\subsubsection{Cosine similarity}
+
+Unlike in the encodings of images, necessary information to decode the data is \textit{not} stored in the encodings of words.
+Natural language is encoded in arbitrary sequences of symbols and provides no useful information to the model regarding the relationships that may exist between various concepts.
+One method that is effectively captures the semantic meaning behind the various concepts are the word embeddings.
+
+Simple approach to extending the word embedding system to a sentence is simply performing some aggregation operation on each of the embeddings.
+We used the summation function aggregation on the words of the sentence, after removing stopwords.
+Stopwords list is used from the NLTK library.
+Word embedding used is the English version of the FastText word embeddings \citep{joulin2016fasttext}.
+
 \subsection{SVM}

### 2017-06-05 paper/main.tex
Part of deep learning explanation added

@@
 \usepackage{tabularx}
+\usepackage{listings}
+\lstset{
+   breaklines=true,
+   basicstyle=\ttfamily}
 
@@
 \begin{center}
-\begin{tabularx}{0.45\textwidth}{llr}
+	\resizebox{\columnwidth}{!}{
+\begin{tabularx}{\columnwidth}{llr}
 \toprule
@@
 \midrule
-4 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
-  & \multicolumn{1}{m{6cm}}{In May 2010, the troops attempted to invade
+4 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
+  & \multicolumn{1}{m{6.5cm}}{In May 2010, the troops attempted to invade
 Kabul.} \\
-& \multicolumn{1}{m{6cm}}{The US army invaded Kabul on May 7th last
+& \multicolumn{1}{m{6.5cm}}{The US army invaded Kabul on May 7th last
 year, 2010.} \\
 \midrule
-3 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
+3 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
   & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
@@
 \midrule
-2 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
+2 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
 &They flew out of the nest in groups. \\
@@
 \midrule
-1 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
+1 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
 & The woman is playing the violin.\\
@@
 0 & \textit{The two sentences are completely dissimilar.}\\
-  & \multicolumn{1}{m{6cm}}{John went horse back riding at dawn with a whole group of friends.}\\
-  & \multicolumn{1}{m{6cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
+  & \multicolumn{1}{m{6.5cm}}{John went horse back riding at dawn with a whole group of friends.}\\
+  & \multicolumn{1}{m{6.5cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
 \bottomrule
 \end{tabularx}
+}
 \end{center}
@@
 
-One of the simplest approaches to measuring semantic similarity is the Jaccard rindex.
+One of the simplest approaches to measuring semantic similarity is the Jaccard index.
 It is a statistic used for comparing similarity and diversity of finite sets.
@@
 
-Simple approach to extending the word embedding system to a sentence is simply performing some aggregation operation on each of the embeddings.
+Simple approach to extending the word embedding system to a sentence is simply performing some aggregation operation on each of the embeddings and then performing the cosine similarity measure.
 We used the summation function aggregation on the words of the sentence, after removing stopwords.
@@
 
+\subsubsection{Motivation}
+Although feature engineering measures are able to provide an approximate measure of similarity, it is trivial to reverse-engineer adversarial examples that completely break the system.
+Consider, for example, two following sentences:
+
+\begin{lstlisting}
+Compared to me, he's heavier.
+I have less kilograms than him.
+\end{lstlisting}
+
+Although their Jaccard index is zero (they share no common words), their semantic similarity is high.
+
+Similar examples can easily be generated by replacing any specific word with its definition:
+\begin{lstlisting}
+Backpropagation.
+Reverse-mode automatic differentiation.
+\end{lstlisting}
+
+Concepts on a such high level of abstraction easily break even the strongest systems since new word definitions could be defined arbitrarily.
+No known models are able to generalize well to strong adversarial examples in an arbitrary language. 
+Such high level generalization could be considered as actual \textit{understanding} of the natural language, which is an AI-complete problem.
+
+A step forward to a model which could provide strong generalization capabilities is the one that has little to none domain specific knowledge embedded inside it and one which has no feature engineering.
+
+One of such models are deep neural networks.
+
+In this paper, one specific instance of such a model is explored: character-level language modelling with deep recurrent neural networks.
+
+\subsubsection{LSTM networks}
+
+Other than human brains, No known models are able to generalize well to such
+
+Although word embeddings are, in theory, more robust to simple word changes,
+
 Deep neural networks have been shown to perform well on language modelling tasks
 (insert citation). 
-Use of highly non-linear functions in the models allows for learning of information on a high level of abstraction.
-Semantic textual similarity bla bla idem spavat sutra ovo dovrsim.
+The non-linear nature of neural networks allows them to model concepts on a high level of abstraction.
+
+Semantic textual similarity exactly that: modelling similarity between concepts that are not shallow, like numbe
 

### 2017-06-05 paper/main.tex
latex updated

--- 2017-06-04 paper/main.tex
@@
 \usepackage{subcaption}
-\usepackage{tabularx}
 
@@
 
-\begin{table}[h!]
+\begin{table*}
 \caption{STS score description, adapted from~\citep{agirre2016semeval}}
@@
 \begin{center}
-\begin{tabularx}{0.45\textwidth}{llr}
+\begin{tabular}{llr}
 \toprule
@@
 \midrule
-4 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
-  & \multicolumn{1}{m{6cm}}{In May 2010, the troops attempted to invade
-Kabul.} \\
-& \multicolumn{1}{m{6cm}}{The US army invaded Kabul on May 7th last
-year, 2010.} \\
+4 & \textit{The two sentences are mostly equivalent, but some unimportant details differ}\\
+& In May 2010, the troops attempted to invade
+Kabul. \\
+& The US army invaded Kabul on May 7th last
+year, 2010.\\
 \midrule
-3 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
-  & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
-suspect.}\\
+3 & \textit{The two sentences are roughly equivalent, but some important information differs/missing.}\\
+& John said he is considered a witness but not a
+suspect.\\
 & ``He is not a suspect anymore.'' John said.\\
 \midrule
-2 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
+2 & \textit{The two sentences are not equivalent, but share some details.} \\
 &They flew out of the nest in groups. \\
@@
 \midrule
-1 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
+1 & \textit{The two sentences are not equivalent, but are on the same topic.}\\
 & The woman is playing the violin.\\
@@
 0 & \textit{The two sentences are completely dissimilar.}\\
-  & \multicolumn{1}{m{6cm}}{John went horse back riding at dawn with a whole group of friends.}\\
-  & \multicolumn{1}{m{6cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
+& John went horse back riding at dawn with a whole group of friends.\\
+& Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.\\
 \bottomrule
-\end{tabularx}
+\end{tabular}
 \end{center}
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[t!]
     \centering
@@
 
-\begin{figure*}[h!]
+\begin{figure*}
 \begin{center}
@@
 score & 0.5898 & 0.7795 \\
+std & 0.011393 & 0.006625 \\
 \bottomrule
@@
 \subsection{Deep learning}
-
-Deep neural networks have been shown to perform well on language modelling tasks
-(insert citation).
-Use of highly non-linear functions in the models allows for learning of information on a high level of abstraction.
-Semantic textual similarity bla bla idem spavat sutra ovo dovrsim.
-
-\subsection{All results}
-
-Table \ref{tab:all_results} shows the performance of all tested models.
-Bla bla we draw some conclusion from this?
-\begin{table}
-	\caption{Model comparison}
-\label{tab:all_results}
-\begin{center}
-\begin{tabular}{ccc}
-\toprule
-Model & Pearson coefficient \\
-\midrule
-Jaccard similarity & 0.73 \\
-Cosine baseline & 0.73 \\
-LSTM & 0.59 \\
-SVM & 0.78 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
 

### 2017-06-05 paper/main.tex
Added deep learning in paper

--- 2017-06-05 paper/main.tex
@@
 
-One of such models are deep neural networks.
+One other way to tackle the problem is from bottom up: by modelling compositions of constituents of the language on various levels of abstractions: characters, words, sentences.
+
+Although vector embeddings seem like a natural way to capture meaning on a word level, capturing composition of words within a sentence is a more difficult problem.
+Summation as an aggregation function doesn't effectively model semantic composition within a sentence.
+It is also invariant to word order within a sentence, while the meaning isn't.
+For example, changing the position of a negation within a sentence changes the meaning, but not the sum.
+
+All of those ideas and problems are a motivation for applying a deep neural network on this task.
 
@@
 
-\subsubsection{LSTM networks}
+\subsubsection{Recurrent networks}
 
-Other than human brains, No known models are able to generalize well to such
+The idea is to apply unsupervised learning with recurrent neural networks to learn a useful model of the language, which could then be used on the semantic similarity problem.
+In other words, we trained the model on a large corpus of text and used transfer learning to use the same trained network on STS.
 
-Although word embeddings are, in theory, more robust to simple word changes,
+We trained a Long short-term memory network \citep{LSTM} on a task of predicting the next character in the complete NLTK Gutenberg corpus. 
+We used a two-layer LSTM cell with 512 size of memory and cross entropy loss on the softmax outputs.
+It is possible to generate arbitrary text with such a trained LSTM network by sampling from the predicted character probability distribution and feeding back that sample to the network.
+One such sample on a trained network is shown in the \ref{tab:sample_output}.
 
-Deep neural networks have been shown to perform well on language modelling tasks
-(insert citation). 
-The non-linear nature of neural networks allows them to model concepts on a high level of abstraction.
+\begin{table}
+\caption{Sample network output}
+\label{tab:sample_output}
+\begin{center}
+\begin{tabular}{c}
+\toprule
+Sample output of the LSTM \\
+\midrule
+\multicolumn{1}{m{6.5cm}}{Little Joe Otter had had no honour pointing to help him so that they
+heard of his own fish.  And it was for the beginning of the Little Fir Tree and
+scrambled to the sheep. 
+Come, let us go and let the right out of this tree, and the LORD had chanced
+to the next tabernacce of Mischians, when the king was carried the short of the children of the Green.}
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
-Semantic textual similarity exactly that: modelling similarity between concepts that are not shallow, like numbe
+Since the network learns an internal representation of the natural language, it is possible to extract the final hidden state of the network and use it as an embedding of the sentence that was fed in as the input.
+To put it in perspective, LSTM here serves as an aggregation function for the one-hot embeddings of the characters.
+It is not invariant to position like the summation, but it doesn't perform well compared to our SVM model. 
+Figure \ref{fig:lstm_corr.png} shows the joint plot of LSTM network.
 
@@
 Table \ref{tab:all_results} shows the performance of all tested models.
-Bla bla we draw some conclusion from this?
+SVM greatly outperforms all other models.
+
 \begin{table}
@@
 Cosine baseline & 0.73 \\
-LSTM & 0.59 \\
+LSTM & 0.61 \\
 SVM & 0.78 \\
@@
 	\begin{center}
-	\includegraphics[width=\columnwidth]{only_2nd_layer.png}
-	\caption{Depiction of blabla}
-	\label{fig:lstm_2nd_layer}
+	\includegraphics[width=\columnwidth]{lstm_corr.png}
+	\caption{Depiction of assigned similarity measure by LSTM language model}
+	\label{fig:lstm_corr}
 	\end{center}
@@
 
+Deep neural network approach could be improved by modelling the language on various levels, not just character levels. 
+Word embeddings and other network architectures could be utilized to gain additional performance gains.
+
 Furthermore additional improvement could be made in bigger model search, both hyperparameters and basic model (e.g.

### 2017-06-05 paper/main.tex
Minor fixes

@@
 We used a two-layer LSTM cell with 512 size of memory and cross entropy loss on the softmax outputs.
+25\% of the NLTK corpus was used as test data. Early stopping was used as an implicit method of regularizing.
 It is possible to generate arbitrary text with such a trained LSTM network by sampling from the predicted character probability distribution and feeding back that sample to the network. 
-One such sample on a trained network is shown in the \ref{tab:sample_output}.
+One such sample on a trained network is shown in the \ref{tab:sample_output} and it helps show, subjectively, the level of comprehension LSTM has of the natural language.
 
@@
 To put it in perspective, LSTM here serves as an aggregation function for the one-hot embeddings of the characters. 
-It is not invariant to position like the summation, but it doesn't perform well compared to our SVM model. 
-Figure \ref{fig:lstm_corr.png} shows the joint plot of LSTM network.
+It has the useful property of not being invariant to the position of characters.
+The final state of first sentence is then compared to the final state of the second sentence using the cosine similarity. 
+In the case of a deep network, like the one we used, only the last layer's final state was used.
+Figure \ref{fig:lstm_corr} shows the joint plot of LSTM network.
 

### 2017-06-05 paper/main.tex
Removed one table from paper

@@
 Beginning quote showcases one such sentance pairs, where human annotators rated 2.8 semantically similar on 0-5 scale, where 0 mean full semantic dissimilarity, and 5 full equivalence with various shades in between.
-Full score semantics you can read in table~\ref{tab:sts-score}
+% Full score semantics you can read in table~\ref{tab:sts-score}
 STS developed system and techniqus have many further applications, transfer learning of sorts.
@@
 
-\begin{table}[h!]
-\caption{STS score description, adapted from~\citep{agirre2016semeval}}
-\label{tab:sts-score}
-\begin{center}
-	\resizebox{\columnwidth}{!}{
-\begin{tabularx}{\columnwidth}{llr}
-\toprule
-Score & Explanation\\
-\midrule
-5 & \textit{Two sentences are completely equivalent}\\
-& The bird is bathing in the sink\\
-& Birdie is washing itself in the water basin.\\
-\midrule
-4 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
-  & \multicolumn{1}{m{6.5cm}}{In May 2010, the troops attempted to invade
-Kabul.} \\
-& \multicolumn{1}{m{6.5cm}}{The US army invaded Kabul on May 7th last
-year, 2010.} \\
-\midrule
-3 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
-  & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
-suspect.}\\
-& ``He is not a suspect anymore.'' John said.\\
-\midrule
-2 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
-&They flew out of the nest in groups. \\
-&They flew into the nest together. \\
-\midrule
-1 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
-& The woman is playing the violin.\\
-& The young lady enjoys listening to the guitar.\\
-\midrule
-0 & \textit{The two sentences are completely dissimilar.}\\
-  & \multicolumn{1}{m{6.5cm}}{John went horse back riding at dawn with a whole group of friends.}\\
-  & \multicolumn{1}{m{6.5cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
-\bottomrule
-\end{tabularx}
-}
-\end{center}
-\end{table*}
-
+% \begin{table}[h!]
+% \caption{STS score description, adapted from~\citep{agirre2016semeval}}
+% \label{tab:sts-score}
+% \begin{center}
+% 	\resizebox{\columnwidth}{!}{
+% \begin{tabularx}{\columnwidth}{llr}
+% \toprule
+% Score & Explanation\\
+% \midrule
+% 5 & \textit{Two sentences are completely equivalent}\\
+% & The bird is bathing in the sink\\
+% & Birdie is washing itself in the water basin.\\
+% \midrule
+% 4 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
+%   & \multicolumn{1}{m{6.5cm}}{In May 2010, the troops attempted to invade
+% Kabul.} \\
+% & \multicolumn{1}{m{6.5cm}}{The US army invaded Kabul on May 7th last
+% year, 2010.} \\
+% \midrule
+% 3 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
+%   & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
+% suspect.}\\
+% & ``He is not a suspect anymore.'' John said.\\
+% \midrule
+% 2 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
+% &They flew out of the nest in groups. \\
+% &They flew into the nest together. \\
+% \midrule
+% 1 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
+% & The woman is playing the violin.\\
+% & The young lady enjoys listening to the guitar.\\
+% \midrule
+% 0 & \textit{The two sentences are completely dissimilar.}\\
+%   & \multicolumn{1}{m{6.5cm}}{John went horse back riding at dawn with a whole group of friends.}\\
+%   & \multicolumn{1}{m{6.5cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
+% \bottomrule
+% \end{tabularx}
+% }
+% \end{center}
+% \end{table*}
+% 
 \section{Related work}

### 2017-07-06 paper/main.tex
Fixed almost all mandatory fixes

@@
 
-\title{Semantic Textual Similarity using deep learning}
+\title{Semantic Textual Similarity Using SVM and Deep Learning}
 
@@
 Unska 3, 10000 Zagreb, Croatia\\
-\texttt{\{bruno.gavranovic,neven.miculinic,stipan.mikulic\}@fer.hr}\\
+\texttt{\{bruno.gavranovic, neven.miculinic, stipan.mikulic\}@fer.hr}\\
 }
@@
 % 
-\section{Related work}
+\section{Related Work}
 
@@
 
-\section{Extent of the paper}
+\section{Extent of the Paper}
 
@@
 
-\subsection{Data analysis}
+\subsection{Data Analysis}
 We used dataset from SemEval 2017
@@
 \begin{table}
-\caption{Stats about lengths of sentences}
+\caption{Stats about lengths of sentences.}
 \label{tab:narrow-table}
@@
 %\end{figure}
-\subsection{Baseline models}
+\subsection{Baseline Models}
 
@@
 
-\subsubsection{Jaccard similarity baseline}
+\subsubsection{Jaccard Similarity Baseline}
 
@@
 
-\subsubsection{Cosine similarity}
+\subsubsection{Cosine Similarity}
 
@@
 \begin{table}
-\caption{Averaged scores}
+\caption{Averaged scores.}
 \label{tab:narrow-table}
@@
 
-\subsection{Deep learning}
+\subsection{Deep Learning}
 
@@
 
-\subsubsection{Recurrent networks}
+\subsubsection{Recurrent Networks}
 
@@
 \begin{table}
-\caption{Sample network output}
+\caption{Sample network output.}
 \label{tab:sample_output}
@@
 
-\subsection{All results}
+\subsection{All Results}
 
@@
 \begin{table}
-	\caption{Model comparison}
+	\caption{Model comparison.}
 \label{tab:all_results}
